Fractional Decomposition Tree Algorithm: A tool for studying the integrality gap of Integer Programs

### 2020-06-12 FDT2020/main-section.tex
arxiv version

\paragraph{FDT on randomly generated instances of TAP.}
Recall that in TAP we are given a tree $T=(V,E)$, and a set of links $L$ between vertices in $V$ and costs $c\in \mathbb{R}^{L}_{\geq 0}$. A feasible augmentation is $L'\subseteq L$ such that $T+L'$ is 2-edge-connected. In TAP we wish to find the minimum-cost feasible augmentation. The integrality gap of the cut-LP for TAP is defined as 
\begin{equation*}
g(\tap) = \max_{c\in \mathbb{R}^L_{\geq 0}} \frac{\min_{x\in\tap(T,L)} cx}{\min_{x\in\cut(T,L)} cx}.
\end{equation*}

We know $\frac{3}{2}\leq g(\tap)\leq 2$~\cite{FJ81,32gaptap}. Notice that $\min_{x\in \tap(T,L)}cx$ is a binary IP. We ran binary FDT on a set of 264 fractional extreme points of randomly generated instances of $\tap$. \cindy{How big were the trees? And how many candidate links?} Table \ref{tableTAP} shows FDT found solutions better than the integrality-gap lower bound for most instances.

\begin{table}[h!]
	\begin{small}
		\centering
		\begin{tabular}{c c c c c}
			\hline
			& $C\in [1.1,1.2]\;$ & $\;C\in (1.2,1.3]\;$ &
			$\;C\in (1.3,1.4]$ &\; $C\in (1.4,1.5]\;$ \\ \hline
			TAP & $36$ & $66$ & $170$ & $10$\\  \hline \\
		\end{tabular}\caption{The scale factor $C$ for FDT run on 264 randomly generated TAP instances with fractional extreme points: 138 instances have $74$ variables. The rest have $250$.} \label{tableTAP}
	\end{small}

\end{table}

\paragraph{Computational comparison between Christofides' algorithm and FDT for 2EC on Carr-Vempala points.}

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=9cm,scale=1.4]{"fdt-vs-christofides".png}
	\caption{Polyhedral version of Christofides' algorithm vs FDT on all Carr-Vempala points that have 10 vertices on the single cycle formed by fractional edges.}
	\label{fdtvschris}
\end{figure}
\paragraph{FDT for 2EC on Carr-Vempala points.}
We ran FDT for 2EC on 963 fractional extreme points of $\subtour(G)$. We enumerated all (fractional) Carr-Vempala points with $10$ and $12$ vertices. Table \ref{table2EC} shows that again FDT found solutions better than the integrality-gap lower bound for most instances. 
\begin{table}[h!]
	\begin{small}
		\centering
		\begin{tabular}{c c c c c}
			\hline
			& $C\in [1.08,1.11]\;$ & $\;C\in (1.11,1.14]\;$ &
			$\;C\in (1.14,1.17]$ &\; $C\in (1.17,1.2]\;$ \\ \hline
			2EC & $79$ & $201$ & $605$ & $78$ \\ \hline\\
		\end{tabular}	\caption{FDT for $\2ec$ implemented applied to all Carr-Vempala with 10 or 12 vertices. A Carr-Vempala point with $k$ vertices has $\frac{3k}{2}$ edges. Thus, the upper bound provided by Theorem \ref{FDT2EC} is $g(\2ec)^{3k/2}$. The lower bound on $g(\2ec)$ is $\frac{6}{5}$.}
		\label{table2EC}
	\end{small}
\end{table}

>>>>

We compared Dive FDT (Algorithm \ref{FDT-dive}) with feasbility pump \cite{fp1} in the terms of time spent solving LP relaxations and the quality of solution provided by each algorithm. The results are presented in Figure \ref{fpvsfdt}.

\begin{figure}[h!]
\begin{subfigure}{.5\textwidth}
\centering
	\includegraphics[width=8cm,scale=1]{"fpvsfdt".png}
	\caption{Dive FDT vs feasbility pump on the instances of PACE 2019 \cite{PACE} with 200 vertices.}
	\label{fpvsfdt}
	\end{subfigure}
	$\quad\;$
	\begin{subfigure}{.5\textwidth}
	\centering
	\includegraphics[width=8cm,scale=1]{"fjvsfdt".png}
	\caption{FDT vs the 2-approximation for TAP\cite{FJ81} on randomly generated extreme points of the cut-LP.}
	\label{fjvsfdt}
	\end{subfigure}
	\caption{Computational experiments with the FDT algorithm}
	\label{fdtcomp}
\end{figure}
\paragraph{FDT on randomly generated instances of TAP.}
Recall that in TAP we are given a tree $T=(V,E)$, and a set of links $L$ between vertices in $V$ and costs $c\in \mathbb{R}^{L}_{\geq 0}$. A feasible augmentation is $L'\subseteq L$ such that $T+L'$ is 2-edge-connected. In TAP we wish to find the minimum-cost feasible augmentation. The integrality gap of the cut-LP for TAP is defined as 
\begin{equation*}
g(\tap) = \max_{c\in \mathbb{R}^L_{\geq 0}} \frac{\min_{x\in\tap(T,L)} cx}{\min_{x\in\cut(T,L)} cx}.
\end{equation*}

We know $\frac{3}{2}\leq g(\tap)\leq 2$~\cite{FJ81,32gaptap}. Notice that $\min_{x\in \tap(T,L)}cx$ is a binary IP. 
We considered full binary trees with 3 to 7 levels. We assume the set of links to be the set of all pairs of leaves of the trees and we choose the costs of the links uniformly at random.  We then ran binary FDT and chose the solution that is gives the minimum cost. We then compared it with the circulation-based 2-approximation algorithm of Frederickson and Ja'Ja' \cite{FJ81} in Figure \ref{fjvsfdt}.

\begin{table}[h!]
	\begin{small}
		\centering
		\begin{tabular}{c c c c}
			\hline
			& number of edges in $T$ & number of links in $L$ & number of instances $(T,L)$\\ \hline
			 & $6$ & $6$ & $100$ \\ 
			 & $14$ & $28$ & $100$ \\ 
			 & $30$ & $120$ & $100$ \\
			 & $62$ & $496$ & $100$ \\ 
			  & $126$ & $2016$ & $50$ \\  \hline \\
		\end{tabular}\caption{Summary of the randomly generated instances of TAP.} \label{tableTAP}
	\end{small}
\end{table}
For all these 450 fractional extreme points of the cut-LP our experiments showed that the integrality gap $g(\tap)$ is at most $\frac{3}{2}$ when restricted to the instances considered. In fact, only for one instance the upper bound we provided by our experiment is at $\frac{3}{2}$.  For 425 instances, the upper bound is $\frac{4}{3}$, for 22 instances it is $\frac{6}{5}$, for 2 instances it is $\frac{8}{7}$.
\paragraph{Computational comparison between Christofides' algorithm and FDT for 2EC on Carr-Vempala points.}

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=9cm,scale=1.4]{christofides-vs-fdt.png}
	\caption{Polyhedral version of Christofides' algorithm vs FDT on all Carr-Vempala points that have 10 vertices on the single cycle formed by fractional edges.}
	\label{fdtvschris}
\end{figure}
\paragraph{FDT for 2EC on Carr-Vempala points.}
We ran FDT for 2EC on 963 fractional extreme points of $\subtour(G)$. We enumerated all (fractional) Carr-Vempala points with $10$ and $12$ vertices. Table \ref{table2EC} shows that again FDT found solutions better than the integrality-gap lower bound for most instances. 
\begin{table}[h!]
	\begin{small}
		\centering
		\begin{tabular}{c c c c c}
			\hline
			& $C\in [1.08,1.11]\;$ & $\;C\in (1.11,1.14]\;$ &
			$\;C\in (1.14,1.17]$ &\; $C\in (1.17,1.2]\;$ \\ \hline
			2EC & $79$ & $201$ & $605$ & $78$ \\ \hline\\
		\end{tabular}	\caption{FDT for $\2ec$ implemented applied to all Carr-Vempala with 10 or 12 vertices. A Carr-Vempala point with $k$ vertices has $\frac{3k}{2}$ edges. Thus, the upper bound provided by Theorem \ref{FDT2EC} is $g(\2ec)^{3k/2}$. The lower bound on $g(\2ec)$ is $\frac{6}{5}$.}
		\label{table2EC}
	\end{small}
\end{table}
